---
output:
  html_document: default
  word_document: default
---
# *ASSIGNMENT 2: ANSWER SCRIPT*
**Load the following Packages**
```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(ggpubr)
```
*Importing  and viewing the dataset*
```{r}
Assignment2_dataset <- read_excel(("C:/Users/DELL/Desktop/RAssignment2/Assignment2_dataset.xlsx"))
View(Assignment2_dataset)
```

### **Question 1**
**The missing values are removed ussing the two functios shown below**
```{r}
Assignment2_New <- na.omit(Assignment2_dataset)
```
### OR
```{r}
Assignment2_New <- na.exclude(Assignment2_dataset) 
```


## **Question 2**
*Show relationship between prices and perceptions*
*define values for perception*
*compute the percentage of each perception that characterize our responses to price.*

```{r}
Positive <- ((length(c(Assignment2_New$P[Assignment2_New$P == "Positive"])))/(length(c(Assignment2_New$P))))*100
Negative <- ((length(c(Assignment2_New$P[Assignment2_New$P == "Negative"])))/(length(c(Assignment2_New$P))))*100
Somewhat_Positive <- ((length(c(Assignment2_New$P[Assignment2_New$P == "SP"])))/(length(c(Assignment2_New$P))))*100
Somewhat_Negative <- ((length(c(Assignment2_New$P[Assignment2_New$P == "SN"])))/(length(c(Assignment2_New$P))))*100
No_Response <- ((length(c(Assignment2_New$P[Assignment2_New$P == "NR"])))/(length(c(Assignment2_New$P))))*100
```

**Displaying the percentage of all perceptions**   
*Negative*   
*Somewhat_Negative*   
*Positive*   
*Somewhat_Positive*   
*No_Response*   

### covariation between perception and price
```{r}
ggplot(data = Assignment2_New, mapping = aes(x = price)) +
  geom_freqpoly(mapping = aes(color = P), binwidth = 500)
```

## **Question3**
*Comparing perception change(PC) and diamond quality (cut)*


```{r}
# **covariation between perception change and diamond quality**
Perception_Change <- Assignment2_New$PC
Cut <- Assignment2_New$cut
ggplot(data = Assignment2_New) +
  geom_count(mapping = aes(x = Cut, y = Perception_Change))
```


## **Question 4**
# *A boxplot of all the variables*
```{r}
boxplot(Assignment2_New$carat, Assignment2_New$depth, Assignment2_New$price, Assignment2_New$x, Assignment2_New$y, names=c("carat", "depth","price", "x", "y"), horizontal=TRUE, col="blue", main="Distribution of variables", xlab="Values", ylab="Variables")
```

*OR ,you can generate individual box plots to better view the outliers for each variable.*
*Below is one of such examples*
```{r,eval=FALSE}
boxplot(Assignment2_New$carat)
```

## **Question5**
### **using the IQR method**
*calculate the IQR for price*
```{r,include=FALSE}
IQRprices <- IQR(Assignment2_New$price)
IQRprices
```

*calculate the quartiles for price*
```{r}
quartileprice <- quantile(Assignment2_New$price, probs=c(.25, .75), na.rm = FALSE)
quartileprice
```

*calculate the upper and lower limits*
```{r}
pricelower <- quartileprice[1] - 1.5*IQRprices
priceupper <- quartileprice[2] + 1.5*IQRprices
pricelower
priceupper
```

*filtering out the outliers for price*
```{r}
price_no_outliers <- subset(Assignment2_New, Assignment2_New$price > priceupper | Assignment2_New$price < pricelower)
price_no_outliers
boxplot(price_no_outliers$price)
```

*saving the new dataframe with no outliers*
```{r}
Dataset <- Assignment2_New[Assignment2_New$price < priceupper & Assignment2_New$price > pricelower,]
```


## **Question 6**
** creating a new csv **

```{r}
write.csv(Dataset, file="C:/Users/DELL/Desktop/A94166/Moga_Muzamil.csv")

```

## **Question 7**
```{r, eval=FALSE}
ggplot(Dataset, aes(x=cut, y=depth)) + geom_bar(stat="identity", fill="red") + labs(title="Relationship between the cut and the depth", x="Cut", y="depth")
```

## **Question 8**
*Compare mean, median and mode*
**Calculating the mean and median of the variable "carat" in the diamonds dataset**
```{r, eval=FALSE}
mean(Dataset$carat,  na.rm = TRUE)
median(Dataset$carat,  na.rm = TRUE)
```

#### OR 

*Q-Q plots draw the correlation between a given sample and the normal distribution.*
```{r,eval=FALSE}
ggqqplot(Dataset$carat, ylab = "Diamond carat",
         ggtheme = theme_minimal())
```


### wilcoxon test on carat
```{r,eval=FALSE}
res <- wilcox.test(Dataset$carat, mu = 0) #mu is theoretical mean (0)
```

## **QUESTION 9**
*Compute the variance between three groups; diamond carat, perception change and price and view the groups*


```{r,echo=FALSE}
groups <- Dataset[,c("carat", "PC", "price")] 
groups
sample <- sample_n(groups, 100)
sample

x <- factor(sample$PC)
x
levels(x)
x <- ordered(x, levels=c("Negative","Positive","SP","Positve","NR","SN"))# the order of the levels
group_by(groups, PC) %>% 
  summarise(
    mean=mean(price), 
    sd=sd(price), n=n())
    
ggline(groups, x = "PC", y = "price", 
        add = c("mean_se"),
        order = c("Negative","Positive","Positve","NR","SN","SP"),
        ylab = "Price", xlab="Perception Change", 
        title = "Line plot of the groups")

```

# Compute the varience.
```{r, echo=FALSE}
anova <- aov(price ~ PC, data = groups)
summary(anova)
```

# **Question 9b**
```{r}
TukeyHSD(anova, conf.level = .95)
plot(TukeyHSD(anova, conf.level = .95))
```
