


---
output:
  html_document: default
 
  pdf_document: default
---

  

```{r echo=FALSE}

  
library(ggplot2)
library(tidyverse)
library(dplyr)
library(ggpubr)
```
## Assignment1 
### Population of the data set are the origins 

### Samples are the varieties 

### Observations are the Flavor/aroma, fragrance,mouth feel,aftertaste ,overall ,salt/acid,bitter/sweet ,balance, uniformity ,clean cups

### Research question :Does performance of varieties vary in different districts ?

### Null Hypothesis:performance of varieties varies in different districts

### Alternative Hypothesis:performance of varieties doesnot vary in different districts

## **1.Descriptive Statistics**
```{r, include=FALSE}
library(readxl)

dataset<- read_excel(("C:/Users/DELL/Desktop/Data-Science-/Group Assignment 1_Cup sensory results-CWDr lines (1).xlsx"))

summary(dataset)


```

- Calculate the  mode  of the overall
#### Mode

```{r echo=FALSE}

getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}

# Create the vector (v) with overall.
v <- c(dataset$OVERALL)
# Calculate the mode using the user function.
getmode(v)

```
 variance and Standard deviation  for overall respectively
```{r echo=FALSE}
var(dataset$OVERALL)
sd((dataset$OVERALL))



```

## Summary of performance in the different regions

- Ibanda summary statistics

```{r echo=FALSE}
Ibanda <- dataset[dataset$ORIGIN == "Ibanda",]
View(Ibanda)
summary(Ibanda$OVERALL)
```
- Mityana summary statistics
```{r echo=FALSE}
Mityana <- dataset[dataset$ORIGIN == "Mityana",]
View(Mityana)
summary(Mityana$OVERALL)
```
- Mukono summary statistics
```{r echo=FALSE}
Mukono <- dataset[dataset$ORIGIN == "Mukono",]
View(Mukono)
summary(Mukono$OVERALL)
```



- Visualise the distribution of the overall across the different districts 

- We can see the histogram for Ibanda is positively skewed 
``` {r echo=FALSE}
hist(Ibanda$OVERALL, main = "Ibanda overall variety performance", xlab = "Overall performance", ylab = NULL)
```

- Mityana is negatively skewed 
``` {r echo=FALSE}
hist(Mityana$OVERALL, main = "Mityana overall variety performance", xlab = "Overall performance", ylab = NULL)
```

``` {r echo=FALSE}
hist(Mukono$OVERALL, main = "Mukono overall variety performance", xlab = "Overall performance", ylab = NULL)
```

- The histogram for Mukono is positively skewed 

### **2.Showing the relationships between the variables** 
``` {r echo=FALSE}
ggplot(dataset, aes(x = dataset$"FLAVOR", y = dataset$"FLAVOR")) + 
  geom_point() + labs(title = "Effect of flavor on overall coffee performance", x = "Overall performance", y = "Flavor")
#Looking at effect of flavor on performance per district
ggplot(Ibanda, aes(x = Ibanda$"FLAVOR", y = Ibanda$"FLAVOR")) + 
  geom_point() + labs(title = "Effect of flavor on coffee performance in Ibanda", x = "Overall performance", y = "Flavor")

ggplot(Mityana, aes(x = Mityana$"FLAVOR", y = Mityana$"FLAVOR")) + 
  geom_point() + labs(title = "Effect of flavor on coffee performance in Mityana", x = "Overall performance", y = "Flavor")

ggplot(Mukono, aes(x = Mukono$"FLAVOR", y = Mukono$"FLAVOR")) + 
  geom_point() + labs(title = "Effect of flavor on coffee performance in Mukono", x = "Overall performance", y = "Flavor")
```




## **3. PERFORMANCE BY VARIETY AND DISTRICT**
**Across districts;**
*Compute medians of districts to show performance since Medians are most suitable rather than the means which were skewed*


*IBANDA MEAN*
```{r}
Ibanda_mean <- mean(Ibanda$OVERALL)
string1<-"Ibanda overall mean is"
result1=paste(string1,Ibanda_mean)
print(result1)
```
*MITYANA MEAN*
```{r}
Mityana_mean <- mean(Mityana$OVERALL)
string2<-"Mityana overall mean is"
result2=paste(string2,Mityana_mean)
print(result2)
```
*MUKONO MEAN*
```{r}
Mukono_mean <- mean(Mukono$OVERALL)
string3<-"Mukono overall mean is"
result3=paste(string3,Mukono_mean)
print(result3)
```

```{r}
Ibanda_median <- median(Ibanda$OVERALL)
ibanda1<-"Ibanda overall median is"
result4=paste(ibanda1,Ibanda_median)
print(result4)

Mityana_median <- median(Mityana$OVERALL)
Mityana1<-"Mityana overall median is"
result5=paste(Mityana1,Mityana_median)
print(result5)

Mukono_median <- median(Mukono$OVERALL)
Mukono1<-"Mukono overall median is"
result6=paste(Mukono1,Mukono_median)
print(result6)
```

**Combine the districts, means or medians**
```{r}
Districts <- c("Ibanda", "Mityana", "Mukono")
Means <- c(Ibanda_mean, Mityana_mean, Mukono_mean)
```
*OR*
```{r}
Median <- c(Ibanda_median, Mityana_median, Mukono_median)
```

**Make a dataframe of the districts and means**
```{r}
Districts_means <- data.frame(Districts,Means)
```

**OR: Make a dataframe of the districts and medians**
```{r}
Districts_medians <- data.frame(Districts,Median)
```

*A Plot of the results*

```{r}
ggplot(Districts_means, aes(Districts,Means,group=1)) +
  geom_line() + geom_point() + labs(title = "Mean overall performance of varieties across districts")
```

**OR Using median**
ggplot(Districts_medians, aes(Districts, Median,group=1)) +
  geom_line() + geom_point() + labs(title = "Median overall performance of varieties across districts")

### **Performance of each variety**
*Transform the dataset to subselect varieties individually*
```{r}
KR3 <- dataset[dataset$VARIETY == "KR3",]
KR4 <- dataset[dataset$VARIETY == "KR4",]
KR5 <- dataset[dataset$VARIETY == "KR5",]
KR6 <- dataset[dataset$VARIETY == "KR6",]
KR7 <- dataset[dataset$VARIETY == "KR7",]
```

**Compute overall means of each variety**
```{r}
mean_KR3 <- mean(KR3$OVERALL)
mean_KR4 <- mean(KR4$OVERALL)
mean_KR5 <- mean(KR5$OVERALL)
mean_KR6 <- mean(KR6$OVERALL)
mean_KR7 <- mean(KR7$OVERALL)
```

### OR

**Compute overall medians of each variety**
```{r}
median_KR3 <- median(KR3$OVERALL)
median_KR4 <- median(KR4$OVERALL)
median_KR5 <- median(KR5$OVERALL)
median_KR6 <- median(KR6$OVERALL)
median_KR7 <- median(KR7$OVERALL)
```


**Combine the varieties, means or medians**
```{r}
Varieties <- c("KR3", "KR4", "KR5", "KR6", "KR7")
Variety_means <- c(mean_KR3,mean_KR4,mean_KR5,mean_KR6,mean_KR7)
Variety_median <- c(median_KR3, median_KR4, median_KR5, median_KR6, median_KR7)
```

**Create a dataframe of Varieties and Variety_means**
```{r}
Dataframe_1 <- data.frame(Varieties, Variety_means)
```

### OR
**Create a datframe of Varieties and Variety_median**
```{r}
Dataframe_2 <- data.frame(Varieties, Variety_median)
```

*A Plot of the results*
```{r}
ggplot(Dataframe_1, aes(Varieties, Variety_means, group=1)) +
  geom_line() + geom_point() + labs (title = "Mean overall variety performance")

```

## **4. CENTRAL TENDENCY/DISTRIBUTION OF EACH VARIABLE.**
### Using a boxplot
**Distribution of overall**
```{r}
boxplot(dataset$OVERALL, main = "OVERALL PERFORMANCE OF VARIETIES", ylab = "overall scores")
```
 
### **QUESTION 5: NORMAL DISTRIBUTION TESTS**
*Test the overall variable for normal distribution using the Shapiro-wilk test*
*Null hypothesis based on research question: The overall performance of varieties across* *districts is normally distributed. p-value >= 0.05* 
*Alternative hypothesis: Overall performance is not normally distributed across districts. p-value =<0.05 *
```{r}
shapiro.test(dataset$OVERALL)
```
*W = 0.95674, p-value = 0.01189*

*Distribution can also be visualised using a  Q-Q plots (quantile-quantile plots). *
*Q-Q plots draw the correlation between a given sample and the normal distribution.*



*The data is not normally distributed, it’s recommended to use the non parametric one-sample Wilcoxon rank test.*
          **WILCOXON ONE-SAMPLE TEST**
*The one-sample Wilcoxon signed rank test is a non-parametric alternative to one-sample t-test when the data cannot be assumed to be normally distributed. *
*it is Used to determine whether the median of the sample is equal to a known standard value (i.e. theoretical value).*
**Looking at the article m =0.8**
```{r}
res <- wilcox.test(dataset$OVERALL, mu = 0.8)
res
```


